---
title: "Homework 1 - Report"
author: "3170105743 李政达"
date: "2020/7/6"
output: 
  pdf_document: 
    latex_engine: xelatex

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

The Iowa data set iowa.csv is a toy example that summarises the yield of wheat (bushels per acre) for the state of Iowa between 1930-1962. In addition to yield, year, rainfall and temperature were recorded as the main predictors of yield.

a. First, we need to load the data set into R using the command read.csv() . Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called iowa.df.

**Ans:** We can use help("read.csv") to learn the parameters of the function. The first parameter is the file name, and it may change in different case. I save the file in the data folder, which is in the parent directory of this .rmd file, so I set the first parameter to `"../data/iowa.csv"`. `header = TRUE` means that the first row contains the names of the variables. `sep = ";"` means that the field separator is semicolon.

```{r}
iowa.df <- read.csv("../data/iowa.csv", header=TRUE, sep = ";")
iowa.df
```

b. How many rows and columns does iowa.df have?

```{r}
length(row.names(iowa.df))
length(names(iowa.df))
```

c. What are the names of the columns of iowa.df?

```{r}
names(iowa.df)
```

d. What is the value of row 5, column 7 of iowa.df?

```{r}
iowa.df[5, 7]
```

e. Display the second row of iowa.df in its entirety.

```{r}
iowa.df[2, ]
```



# Question 2

Syntax and class-typing.

a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.

```
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
sum(vector1)
```

**Ans:**

- `c()` can combine values into a vector or list, so `vector1` will be a vector containing `"5"  "12" "7"  "32"`.
- `max()` return the maximum of the input values. The type of input is character, which can be compare. So the command runs correctly and returns the maximum, which is `"7"`.
- `sort()` can sort a vector into ascending order. The type of input is `character`, which can be compare. So the command runs correctly and sorts the vector, and outputs `"12" "32" "5"  "7" `.
- `sum()` can compute the sums of all the input values, but the character type can not be added, so the command will cause an error.

```{r}
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
#sum(vector1) #error
```

b. For the next series of commands, either explain their results, or why they should produce errors.

```
vector2 <- c("5",7,12)
vector2[2] + vector2[3]

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
list4[2]+list4[4]
```

**Ans:** 

- `c()` will combine all kinds of values, but the type of the values may change in order to compatibility. For example, boolean variable will change its type when combined with integer variable, and integer variable will change its type when combined with double variable. In this command, the type of the elements in `vector2` will be character.
- Because the values of `vector2` is character, they can not be added. So the command will cause an error.