---
title: "Homework 3 - Report"
CJKmainfont: SimSun
author: "3170105743 李政达"
date: "2020/7/8"
output:
  pdf_document:
    includes:
      header-includes:
        - \usepackage{xeCJK}
    keep_tex: no
    latex_engine: xelatex

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You will estimate the power-law scaling model, and its uncertainty, using the data alluded to in lecture, available in the file `gmp.dat` from lecture, which contains data for 2006.

```{r}
gmp <- read.table("../data/gmp.dat")
gmp$pop <- round(gmp$gmp / gmp$pcgmp)
```

1. First, plot the data as in lecture, with per capita GMP on the y-axis and population on the x-axis. Add the curve function with the default values provided in lecture. Add two more curves corresponding to $a=0.1$ and $a=0.15$; use the `col` option to give each curve a different color (of your choice).

```{r}
plot(gmp$pcgmp ~ gmp$pop, log = "x",
     xlab = "population", ylab = "per capita GMP",
     main = "US Metropolitan Areas, 2006")
curve(6611*x^(1/8), add = TRUE, col = "blue")
curve(6611*x^(0.1), add = TRUE, col = "red")
curve(6611*x^(0.15), add = TRUE, col = "green")
```

